The Folly of Passwords

### 2014-12-13 paper.tex
First pass done. Needs lots of cutting.

\begin{itemize}
  \item users can sign into desktop machines using the authentication method
    being evaluated;
  \item users can then access their corporate email without needing additionally
    supplying a password to do so;
  \item users are able to sign in to a dummy web service (a simple static page
    behind an authentication layer was used);
  \item services should be able to acquire information about a user for
    authorisation purposes (e.g. determining a user's department from
    a corporate user directory service); and
  \item users should be able to do remote SSH login to development Unix
    servers (this assumes a company where this is a common work task, e.g.
    software development, which is including to exercise the authentication
    further rather than to act as a typical business use case).
\end{itemize}

@@

Some desktop environments (e.g. Gnome, Mac OS X, etc.) provide
a local keychain that can store SSH keys, SSL certificates, etc.
which are unlocked whilst the user is logged in, but inaccessible
otherwise. This already provides a convenience desktop
users, but it relies on the user authenticating with a password
that is then the same password used to update the keychain
(Mac OS X updates the keychain password when the user's main
account password is changed). There maybe be some scope for
further research into whether dekstop keychains can be made
compatible with OTP authentication.

@@

@@

@@

>>>>

\begin{itemize}
  \item \textbf{Desktop Login}: users can sign into desktop machines using the authentication method
    being evaluated;
  \item \textbf{Company Email}: users can then access their corporate email without needing additionally
    supplying a password to do so;
  \item \textbf{Web Login}: users are able to sign in to a dummy web service (a simple static page
    behind an authentication layer was used);
  \item \textbf{SSH Login}: users should be able to do remote SSH login to development Unix
    servers (this assumes a company where this is a common work task, e.g.
    software development, which is including to exercise the authentication
    further rather than to act as a typical business use case).
\end{itemize}

@@

LinOTP also provides centralised web-based identity approaches
such as OpenID and SAML. A company-wide identity provider would
not only allow Single Sign-On within a series of web-based
internal applications, but opens up the possibility of employees
being able to use their company identity elsewhere. One example
is if the company wishes to purchase access to online resources
(online books, journals or any other cloud-based service), they
could make agreement with the provider to allow employees to log in
via OpenID or SAML. Universities already use this model for access to online
journals and academic materials.

Some desktop environments (e.g. Gnome, Mac OS X) provide
a local keychain that can store SSH keys, SSL certificates, etc.
which are unlocked whilst the user is logged in, but inaccessible
otherwise. This already provides a convenience desktop
users, but it relies on the user authenticating with a password
that is then the same password used to update the keychain
(Mac OS X updates the keychain password when the user's main
account password is changed). There maybe be some scope for
further research into whether dekstop keychains can be made
compatible with OTP authentication.

@@

An Apache module is also available
for both the Yubikey and also generic OTP logins. Whilst having multiple
login points all asking for a new OTP would be an inconvenience when using a mobile
app, this is less of a problem for the YubiKey where the user needs only press
a single button each time. This might allow for internal applications each to
implement their own OTP support and for them to be managed separately, whereas
the OTP mobile app approach would be better with true Single Sign-On.

@@

Table~\ref{table:overview} summarises the approaches tried on our test network
and discussed in chapter~\ref{chapter:alternatives}. In some cases, there
are a few alternatives listed where some aspects of more than one approach
were explored. These are by no means a definitive list of all options available
to individuals and organisations and some of the technologies and systems
not evaluated thoroughly are mentioned in passing throughout
chapter~\ref{chapter:alternatives}.

\begin{table}[h]
  \begin{tabular}{|l|l|l|l|l|} \hline
                            & \textbf{OTP Apps} & \textbf{pam\_usb} & \textbf{OpenPGP smartcard} & \textbf{YubiKey} \\ \hline
    \textbf{Desktop Login}  & pam\_oath        & pam\_usb                    & pam\_poldi & pam\_oath, pam\_yubico  \\ \hline
    \textbf{Company Email}  & LinOTP + Kerberos & X.509 certificate on device & Scute      & LinOTP + Kerberos, static password + OpenLDAP   \\ \hline
    \textbf{Web Login}      & OpenID, SAML     & X.509 certificate on device & Scute      & Apache YubiKey/OTP modules, static password + OpenLDAP  \\ \hline
    \textbf{SSH Login}      & Gnome keyring    & SSH key stored on device    & gpg-agent  &  \\ \hline
  \end{tabular}
  \label{table:overview}
\end{table}

Overall, YubiKeys seem to provide the best balance between application support,
usability and compatibility. OpenPGP and other smartcards are perhaps also
a good approach where security is of greater concern and there is some appeal
in terms of usability given users are already used to keeping credit cards
safe and protecting them with short PINs. Their requirement for specialised
card readers and the fact it is still best practice to protect them with memorised
information means significant monetary and effort costs that may be acceptable
to trade off for the additional security, but this must be considered.

LinOTP is a promising enterprise solution that is compatible with Windows
domain setups common in a lot of companies. it appears to provide the most
flexibility in terms of application support.

Perhaps the most important YubiKey feature in practical terms is the ability
to store a static password alongside its OTP behaviour. This enables its
use in \emph{existing} corporate networks where users have a single username-password
pair and all systems authenticate against a directory service such as Active Directory
or OpenLDAP. To set this up, security departments need only include synchronsing
users' passwords to their YubiKeys as part of account creation. Users' passwords
can thus become much stronger and more random without the risk of forgetting or
users resorting to writing them down.

The static password feature is entirely compatible with every system that already
supports password authentication, so the only additional costs are purchasing
the hardware and ensuring users are given quick replacements with new passwords
when they report their existing key lost or stolen. Users are more aware of the
need to report lost or stolen physical items (company ID card, house keys, credit cards, etc.)
whereas the evidence in the literature around passwords suggests users disclose
passwords far more readily (whether deliberately or inadvertantly).

Thus we can recommend coporations are able to evaluate the hardware costs
and consider YubiKeys as a ``drop in'' replacement for users memorising
passwords with few infrastructure changes. This strategy even lends itself
to the possiblity of being run as a trial with one team or department. Over time,
more structural changes such as introducing LinOTP could be considered to
reap some of the benefits OTP and challenge-response bring (e.g. protection
from replay attacks, reduced exposure to brute-force guessing).

It may be possible to run the static password trial alongside introducing OTP
for some applications only. The YubiKey supports generating either token
by using a short versus a long button press, but the usability of that has
not been demonstrated.

@@

The following questions are open and are possibilities for further research:

\begin{itemize}
\item What is the effectiveness and what potential issues are there around
  a real deployment of LinOTP within a business? An empirical study in a
  real world setting would give some valuable insights.
\item What is the best approach (if any) to storing keys and certificates on
  a USB device, keeping them secure and allowing them to be used in onward,
  automatic authentication against other services after login? Can a user's
  own USB storage device (already commonplace in many offices) be thus set
  up as an authentication token without the need for specialist hardware
  such as YubiKeys?
\item Today, how easily can an IT or security department set up users with
  smartcards to provide a Single Sign-On experience across all company
  systems? A real world case study would be really useful to bring together
  all the individual projects' documentation and prove an end-to-end deployment
  of them.
\item Are users happy to treat smartcards such as OpenPGP cards within the
  same guidelines already followed for credit cards?
  \item What real world case studies exist for enterprise deployments of YubiKeys?
    Yubico claim multiple corporate customers on their website, but have any
    been studied in the academic literature?
\item Many of the mechanisms discussed are seeing large-scale
  adoption as a second factor in addition to passwords. However, What are the
  security implications of relying on any password alternative as a single factor?
  Do any of them end up being more vulnerable than a traditional, memorised
  password when used in isolation?
\end{itemize}
